Energy and the CES production function

### 2015-08-24 Articles/Article_EcolEcon_2015/EcolEcon_2015.Rnw
Move alpha resampling graph.
New version of SSE vs. lambda graph.

<<model_points_style>>=
model_points_style <- function(p) {
  p + 
  scale_colour_manual(values = c("gray70", "black")) +
  scale_size_manual(values = c(5, 3,3,3,3,3,3)) +
  # scale_alpha_manual(values = c(.2, .6,.6,.6,.6, .6, .6)) +
  scale_fill_manual(values = c(scales::alpha("gray50", .2), rep("transparent", 6))) +
  scale_shape_manual(values = c(21, 21, 23:27)) +
  guides( 
    col = guide_legend(override.aes = aes(size = 5, shape = 21)) )
}
@

@@

  p <- ggplot() +
    geom_point(
      data=dataToPlot, 
      mapping=aes(
        x=Year, 
        y=Residuals, 
        fill=Model, 
        shape=Model, 
        size=Model,
        color=Inputs)
      ) +
    facet_grid(flavor ~ Country) +
    geom_hline(yintercept = 0, colour="gray50", size=0.1) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    xlab("") +
    ylab("r") + 
    xy_theme() +
    labs(color = "Factors of production") # +
    # theme(legend.title=element_blank())

  model_points_style(p)
}
@

@@

@@

p <- ggplot() +
  geom_point(
    data=data_sr_sse, 
    mapping=aes(
      x=lambda, 
      y=sse, 
      fill=Model, 
      shape=Model, 
      size=Model, 
      color=Inputs)) +
  geom_vline(aes(xintercept = lambda), 
             AllCoeffs %>% filter(model=="exp") %>% select(Country, lambda), 
             colour="gray50", size=0.1) +
  geom_hline(aes(yintercept = sse), 
             AllCoeffs %>% filter(model=="exp") %>% select(Country, sse), 
             colour="gray50", size=0.1) +
  facet_grid(flavor ~ Country) +
  scale_x_continuous(breaks=c(0, 0.01, 0.02, 0.03), limits=c(0, 0.035)) +
  xlab(expression(lambda)) +
  scale_y_continuous(limits=c(-0.05, 0.7)) +
  ylab("SSE") + 
  labs(color = "Factors of production") +
  xy_theme() # +
  # theme(legend.title=element_blank())

@@

@
%
Figures~\ref{fig:Factor_shares_graph_PT} and~\ref{fig:Factor_shares_graph_UK}
show factor shares ($\alpha$ values) for CES models for
Portugal and the UK, respectively.
See Table~\ref{tab:CES_abg_equations} for details on the calculation
of factor shares from fitted values of $\delta_1$ and $\delta$.
%
<<Factor_shares_bar_graph_function>>=
fsbg <- function(coun, mod){
  bar_chart_data <- 
    AllCoeffs %>% 
    filter(Country==coun & model==mod) %>% 
    select(Country, model, flavor, energy, nest, CSP, alpha_1, alpha_2, alpha_3) %>%
    mutate(nest =   # Change nest to a nicer-looking nest string
             plyr::mapvalues(nest, 
                             from=c("kl", "kle", "lek", "ekl"), 
                             to=c("kl", "(kl)e", "(le)k", "(ek)l")),
           label = # Add a label column
              paste(CSP, flavor, energy, nest, sep=", "),
           label =  # Add row numbers to the label column
              paste(paste0(1:length(label), ". "), label)
    )
  # Capture the order that we want for the models on the vertical axis.
  order <- bar_chart_data$label
  # Now melt the data
  ### tidyr::spread and tidyr::gather are preferred in most cases now
#   bar_chart_data <- melt(bar_chart_data, 
#                          measure.vars = c("alpha_1", "alpha_2", "alpha_3"), 
#                          value.name = "value", 
#                          variable.name = "variable")
  bar_chart_data <-
    bar_chart_data %>% gather( key=variable, value=value, alpha_1 : alpha_3 )

@@

<<output_elasticities_graph_ternary_2, fig.pos="H", fig.align="center", fig.width=7.5, fig.height=5, fig.cap="Ternary graph of output elasticities~($\\alpha$). This version splits the Unadjusted and Quality-adjusted values on their own graphs. This approach is parallel to the previous figures (historical data, fitted GDP, and residuals.">>=
triPlot(
  data=
    AllCoeffs %>% filter(model == "CES") %>%
    mutate( Model = interaction(nest, CSP, sep = " ", drop=TRUE),
            Inputs = flavor), 
  grid_lines = 5,
  labels = c("alpha[k]", "alpha[l]", "alpha[e]"),
  aes(x = alpha_1, y = alpha_2, z = alpha_3, 
      fill  = Model, 
      shape = Model,
      size = Model
      # color = Inputs
      )
  ) +
  facet_grid(Inputs ~ Country) +

    # model_points_style(p)

  # scale_colour_manual(values = c("gray70", "black")) +
  scale_size_manual(values = c(5, 3,3,3,3,3,3)) +
  # scale_alpha_manual(values = c(.2, .6,.6,.6,.6, .6, .6)) +
  scale_fill_manual(values = c(scales::alpha("gray50", .2), rep("transparent", 6))) +
  scale_shape_manual(values = c(21, 21, 23:27)) +
  guides( 
    col = guide_legend(override.aes = aes(size = 5, shape = 21)) )  
@

@@

<<sigma_graph_3, fig.pos="H", fig.align='center', fig.width=8.5, fig.height=5, out.width="0.9\\textwidth", fig.cap="Elasticities of substitution~($\\sigma$) with factors of production on the vertical facet.">>=
breaks_labels = c(0.01, 0.1, 1, 10, 100)
p <- 
  ggplot(
    AllCoeffs %>% 
      gather(type, value, matches("sigma")) %>% 
      filter(!is.na(value)) %>% 
      filter(! type %in% c("sigma", "sigma_1")) %>%
      mutate(Model = interaction(nest, CSP, sep = " ", drop=TRUE),
             Inputs = flavor
             )
    ) + 
  geom_point(aes(x=value, y=type, 
                 shape = Model, fill=Model, size=Model,
                 color=Inputs
                 )
             #position = position_jitter(height=0.2, width=0)
             ) + 
  facet_grid( Inputs ~ Country) + 
  scale_x_log10(limits = c(0.001,2000), breaks=breaks_labels, labels=breaks_labels) +
  xlab("") +
  geom_vline(xintercept = 1, color="gray50", size=0.2) +
  ylab("") +
  labs(color = "Factors of production") +
  xy_theme()

>>>>

<<model_points_style>>=
model_points_style <- function(p) {
  p + 
  scale_colour_manual(values = c("gray70", "black")) +
  scale_size_manual(values = c(5, 3,3,3,3,3,3)) +
  # scale_alpha_manual(values = c(.2, .6,.6,.6,.6, .6, .6)) +
  scale_fill_manual(values = c(scales::alpha("gray50", .2), rep("transparent", 6))) +
  scale_shape_manual(values = c(21, 21, 23:27)) +
  guides( 
    col = guide_legend(override.aes = aes(size = 5, shape = 21)) )
}

model_points_style_quad <- function(p) {
  p + 
  # scale_colour_manual(values = c("gray70", "black")) +
  scale_size_manual(values = c(5, 3,3,3,3,3,3)) +
  # scale_alpha_manual(values = c(.2, .6,.6,.6,.6, .6, .6)) +
  scale_fill_manual(values = c(scales::alpha("gray50", .2), rep("transparent", 6))) +
  scale_shape_manual(values = c(21, 21, 23:27)) +
  guides( 
    col = guide_legend(override.aes = aes(size = 5, shape = 21)) 
    ) + 
  theme(legend.title=element_blank())
}
@

@@

  p <- ggplot() +
    geom_point(
      data=dataToPlot, 
      mapping=aes(
        x=Year, 
        y=Residuals, 
        fill=Model, 
        shape=Model, 
        size=Model #,
        # color=Inputs
        )
      ) +
    facet_grid(flavor ~ Country) +
    geom_hline(yintercept = 0, colour="gray50", size=0.1) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    xlab("") +
    ylab("r") + 
    xy_theme() +
    labs(color = "Factors of production") # +
    # theme(legend.title=element_blank())

  model_points_style_quad(p)
}
@

@@

%++++++++++++++++++++++++++++++
\subsection{SSE and Solow residual} 
\label{sec:SSE_and_SR_graphs}
%++++++++++++++++++++++++++++++

@@

p <- ggplot() +
  geom_point(
    data=data_sr_sse, 
    mapping=aes(
      x=lambda, 
      y=sse, 
      fill=Model, 
      shape=Model, 
      size=Model, 
      color=Inputs
      )) +
  geom_vline(aes(xintercept = lambda), 
             AllCoeffs %>% filter(model=="exp") %>% select(Country, lambda), 
             colour="gray50", size=0.1) +
  geom_hline(aes(yintercept = sse), 
             AllCoeffs %>% filter(model=="exp") %>% select(Country, sse), 
             colour="gray50", size=0.1) +
  facet_grid(. ~ Country) +
  scale_x_continuous(breaks=c(0, 0.01, 0.02, 0.03), limits=c(0, 0.035)) +
  xlab(expression(lambda)) +
  scale_y_continuous(limits=c(-0.05, 0.7)) +
  ylab("SSE") + 
  labs(color = "Factors of production") +
  xy_theme() # +
  # theme(legend.title=element_blank())

@@

<<SRsse_CES_graph_3, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Solow Residuals~($\\lambda$) and mean squared error~($sse$). $\\lambda$ and $sse$ values for the reference (exponential-only) model are shown as gridlines. This version uses our proposed new style.">>=
data_sr_sse <- 
  AllCoeffs %>% 
  filter(model=="CES") %>% 
  mutate(nest = 
           plyr::mapvalues(
             nest, 
             from=c("kl", "kle", "lek", "ekl"), 
             to=c("kl", "(kl)e", "(le)k", "(ek)l")
             ),
         Model = interaction(nest, CSP, sep = " ", drop=TRUE),
         Inputs = flavor
  )

p <- ggplot() +
  geom_point(
    data=data_sr_sse, 
    mapping=aes(
      x=lambda, 
      y=sse, 
      fill=Model, 
      shape=Model, 
      size=Model #, 
      # color=Inputs
      )) +
  geom_vline(aes(xintercept = lambda), 
             AllCoeffs %>% filter(model=="exp") %>% select(Country, lambda), 
             colour="gray50", size=0.1) +
  geom_hline(aes(yintercept = sse), 
             AllCoeffs %>% filter(model=="exp") %>% select(Country, sse), 
             colour="gray50", size=0.1) +
  facet_grid(flavor ~ Country) +
  scale_x_continuous(breaks=c(0, 0.01, 0.02, 0.03), limits=c(0, 0.035)) +
  xlab(expression(lambda)) +
  scale_y_continuous(limits=c(-0.05, 0.7)) +
  ylab("SSE") + 
  labs(color = "Factors of production") +
  xy_theme() # +
  # theme(legend.title=element_blank())

# Add our style
model_points_style_quad(p)

Figures~\ref{fig:Factor_shares_graph_PT} and~\ref{fig:Factor_shares_graph_UK}
show factor shares ($\alpha$ values) for CES models for
Portugal and the UK, respectively.
See Table~\ref{tab:CES_abg_equations} for details on the calculation
of factor shares from fitted values of $\delta_1$ and $\delta$.
%
<<Factor_shares_bar_graph_function>>=
fsbg <- function(coun, mod){
  bar_chart_data <- 
    AllCoeffs %>% 
    filter(Country==coun & model==mod) %>% 
    select(Country, model, flavor, energy, nest, CSP, alpha_1, alpha_2, alpha_3) %>%
    mutate(nest =   # Change nest to a nicer-looking nest string
             plyr::mapvalues(nest, 
                             from=c("kl", "kle", "lek", "ekl"), 
                             to=c("kl", "(kl)e", "(le)k", "(ek)l")),
           label = # Add a label column
              paste(CSP, flavor, energy, nest, sep=", "),
           label =  # Add row numbers to the label column
              paste(paste0(1:length(label), ". "), label)
    )
  # Capture the order that we want for the models on the vertical axis.
  order <- bar_chart_data$label
  # Now melt the data
  ### tidyr::spread and tidyr::gather are preferred in most cases now
#   bar_chart_data <- melt(bar_chart_data, 
#                          measure.vars = c("alpha_1", "alpha_2", "alpha_3"), 
#                          value.name = "value", 
#                          variable.name = "variable")
  bar_chart_data <-
    bar_chart_data %>% gather( key=variable, value=value, alpha_1 : alpha_3 )

@@

<<output_elasticities_graph_ternary_2, fig.pos="H", fig.align="center", fig.width=7.5, fig.height=5, fig.cap="Ternary graph of output elasticities~($\\alpha$). This version splits the Unadjusted and Quality-adjusted values on their own graphs. This approach is parallel to the previous figures (historical data, fitted GDP, and residuals.">>=
p <- triPlot(
  data=
    AllCoeffs %>% filter(model == "CES") %>%
    mutate( Model = interaction(nest, CSP, sep = " ", drop=TRUE),
            Inputs = flavor), 
  grid_lines = 5,
  labels = c("alpha[k]", "alpha[l]", "alpha[e]"),
  aes(x = alpha_1, y = alpha_2, z = alpha_3, 
      fill  = Model, 
      shape = Model,
      size = Model
      # color = Inputs
      )
  ) +
  facet_grid(Inputs ~ Country)

  model_points_style_quad(p)
@

@@

<<sigma_graph_3, fig.pos="H", fig.align='center', fig.width=8.5, fig.height=5, out.width="0.9\\textwidth", fig.cap="Elasticities of substitution~($\\sigma$) with factors of production on the vertical facet.">>=
breaks_labels = c(0.01, 0.1, 1, 10, 100)
p <- 
  ggplot(
    AllCoeffs %>% 
      gather(type, value, matches("sigma")) %>% 
      filter(!is.na(value)) %>% 
      filter(! type %in% c("sigma", "sigma_1")) %>%
      mutate(Model = interaction(nest, CSP, sep = " ", drop=TRUE),
             Inputs = flavor
             )
    ) + 
  geom_point(aes(x=value, y=type, 
                 shape = Model, fill=Model, size=Model#,
                 # color=Inputs
                 )
             #position = position_jitter(height=0.2, width=0)
             ) + 
  facet_grid( Inputs ~ Country) + 
  scale_x_log10(limits = c(0.001,2000), breaks=breaks_labels, labels=breaks_labels) +
  xlab("") +
  geom_vline(xintercept = 1, color="gray50", size=0.2) +
  ylab("") +
#   labs(color = "Factors of production") +
  xy_theme()

model_points_style_quad(p)
@
